---
title: "Untitled"
author: "Patrick Corbett"
date: "12/23/2022"
date_modified: "08/25/2024"
output: html_document
---


**Banner when running App**
```{r}
# Banner when running App
if( exists("app_run")) {
   showModal(modalDialog("Generating Hot Spots", footer=NULL))
}
```


**Set Parameters for running Outside the Rshiny App**
```{r, setup, include=FALSE}
time_start <- Sys.time()
if(! exists("app_run")) {
  #Establish Root Directory
  knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
  #Parameters
 countries="Burundi"
 neighbor_distance= 1700
 
}
dogdensfx = FALSE
```


# Setting Directories
```{r}

country_folder <- paste("Inputs/Country_Files/",countries,sep="")

output_folder <- paste0("Outputs/Country_Outputs/",countries )

folder <- paste0("Outputs/Country_Outputs/",countries, "/Clusters" )

dir.create(folder, recursive = TRUE)

```


# Read in STARC and HDR Data 

```{r}
# Read STARC File
file <- paste0(output_folder,"/STARC_Map.shp")
starc_code_ <- st_read(file)

# Read HDRs csv file (This needs to be manually specified per country)
hdrs = read.csv(paste0("Inputs/HDRs/",countries,"_HDRs.csv" ))

```

# Compute Estimated Dog Pop and Densities

```{r}
# Merge hdrs and starc file
starc_code_ <- starc_code_ %>% filter(st_geometry_type(geometry)=="POLYGON") %>% 
  merge(hdrs, by="strc_cd", all.x=T) 

# Calculate Pop by category (necessary later)
starc_code_2 <- starc_code_ %>% mutate(
                            pop_s1 = ifelse(RUC=="S1", pop_sum, NA),
                            pop_s2 = ifelse(RUC=="S2", pop_sum, NA),
                            pop_s3 = ifelse(RUC=="S3", pop_sum, NA),
                            pop_s4 = ifelse(RUC=="S4", pop_sum, NA),
                            pop_s5 = ifelse(RUC=="S5", pop_sum, NA),
                            # All Dogs
                            Dgs_s1_A = ifelse(RUC=="S1", pop_sum/All_HDR, NA),
                            Dgs_s2_A = ifelse(RUC=="S2", pop_sum/All_HDR, NA),
                            Dgs_s3_A = ifelse(RUC=="S3", pop_sum/All_HDR, NA),
                            Dgs_s4_A = ifelse(RUC=="S4", pop_sum/All_HDR, NA),
                            Dgs_s5_A = ifelse(RUC=="S5", pop_sum/All_HDR, NA),
                            # Free Roaming
                            Dgs_s1_F = ifelse(RUC=="S1", pop_sum/Free_HDR, NA),
                            Dgs_s2_F = ifelse(RUC=="S2", pop_sum/Free_HDR, NA),
                            Dgs_s3_F = ifelse(RUC=="S3", pop_sum/Free_HDR, NA),
                            Dgs_s4_F = ifelse(RUC=="S4", pop_sum/Free_HDR, NA),
                            Dgs_s5_F = ifelse(RUC=="S5", pop_sum/Free_HDR, NA),
                            # ZAMBIA ONLY - dog density fx
                            dg_dns_fxA = 1.59 * exp(0.000244*pp_dnst),
                            dg_dns_fxF = 1.59 * exp(0.000506*pp_dnst),
                            dg_pp_fxA = dg_dns_fxA*ar_sqkm,
                            dg_pp_fxF = dg_dns_fxF*ar_sqkm
                            ) %>% rowwise() %>% 
                            mutate(
              dg_pp_A= sum(Dgs_s1_A, Dgs_s2_A, Dgs_s3_A, Dgs_s4_A,Dgs_s5_A, na.rm=T ),
              dg_pp_F= sum(Dgs_s1_F, Dgs_s2_F, Dgs_s3_F, Dgs_s4_F,Dgs_s5_F, na.rm=T )
                                                        )
starc_code_2 <-starc_code_2 %>% mutate(
  dg_dns_A = round(dg_pp_A/ar_sqkm, digits=3),
  dg_dns_F = round(dg_pp_F/ar_sqkm, digits=3) )
starc_code_2$id <- seq_along(data.frame(starc_code_2%>%st_drop_geometry())[,2])

if(dogdensfx){
  starc_code_2 = starc_code_2 %>% 
    dplyr::select(-dg_pp_A,-dg_pp_F,-dg_dns_A,-dg_dns_F) %>%
    rename(dg_pp_A = dg_pp_fxA, dg_pp_F = dg_pp_fxF,
           dg_dns_A = dg_dns_fxA, dg_dns_F = dg_dns_fxF,)
}
if(!dogdensfx){
  starc_code_2 = starc_code_2 %>% 
    dplyr::select(-dg_pp_fxA,-dg_pp_fxF,-dg_dns_fxA,-dg_dns_fxF)
}


```

# Regional STARC
```{r}

# country1SL <- starc_code_2
# country2GU <- starc_code_2
# country3LI <- starc_code_2
# country4CI <- starc_code_2
# # country5 <- starc_code_2
# 
# starc_code_2 = rbind(country1SL,
#                     country2GU%>%st_transform(st_crs(country1SL)),
#                     country3LI%>%st_transform(st_crs(country1SL)),
#                     country4CI%>%st_transform(st_crs(country1SL)))



# country1 <- starc_code_2
# country2 <- starc_code_2
# 
# country11 = country1%>%dplyr::select(-"shapeName.1",-"shapeISO.1",-"shapeID.1",
#                                      -"shapeGroup.1",-"shapeType.1")
# country22 = country2%>%dplyr::select(-"All_HDR.x",-"Free_HDR.x",-"reg.x",-"All_HDR.y",-"Free_HDR.y",-"reg.y")
# 
# starc_code_2 = rbind(country11,
#                     country22)
# 
# hdrs = read.csv(paste0(country_folder, "/",countries,"_HDRs.csv" ))
# 
# test = starc_code_2[starc_code_2$shapeName=="Bali",]

# countryTL <- starc_code_2
# # countryIN <- starc_code_2
#
# starc_code_2 = rbind(country2,
                    # country1%>%st_transform(st_crs(country2))%>%dplyr::select(names(country2)))

```

# Getis-Ord Statistic Test 

## dnearneigh Function 
**Function for speeding up the calculation of  distance neighbors**
```{r}
function_using_spdep <- function(pts, min, max) {
  nc <- 7L # number of cores
  cores_opt <- set.coresOption(nc)
  mc_opt <- set.mcOption(FALSE)
  cl <- parallel::makeCluster(get.coresOption())
  set.ClusterOption(cl)
  on.exit({
    set.coresOption(cores_opt)
    set.mcOption(mc_opt)
    set.ClusterOption(NULL)
    parallel::stopCluster(cl)
  })

  # do spdep stuff
  dneigh <- dnearneigh(pts, min, max)
  return(dneigh)
}

```

Gi (hexagons are not neighbors to selves): focal analysis 
Gi* (hexagons are neighbors to selves): local analysis 

## Calculate Gi*

```{r}

a=  Sys.time()
# Get all neighbors for the centroids of each hexagon within certain distance (only "neighbor" if centroid is within that distance)
dnb = function_using_spdep(st_coordinates(st_centroid(starc_code_2)), 0, neighbor_distance)
b=Sys.time()
b-a

# Make each point it's own neighbor, necessary for the gi* statistic
dnb_self = include.self(dnb) 

# construct a row-standardised weights matrix. Row standardization involves dividing each neighbor weight for a feature by the sum of all neighbor weights for that feature, and is recommended whenever the distribution of your features is potentially biased due to sampling design or an imposed aggregation scheme. https://rpubs.com/tskam/IS415-Hands-on_Ex07 
dnb_lw <- nb2listw(dnb_self, style = 'B', zero.policy = T)

# Calculate the actual Gi* statistic based on SOME quantatitative variable (dog density, dog frequency, etc) which is also essentially a unstandardized Z score
gi.fixed <- localG(starc_code_2$dg_dns_F , dnb_lw,alternative="greater") # or use dg_pp_F (free roaming dog population...not density)
gi_test_info = data.frame(attr(gi.fixed,"internals")) %>% mutate(Pr.z...E.G.i.. = round(Pr.z...E.G.i.., digits=6), 
                                                                 cluster= attr(gi.fixed,"cluster"))

  
# Standardize the Z score
gi_sig <- spdep::hotspot(gi.fixed, "Pr(z > E(G*i))", cutoff=0.005,droplevels=T)# Room to edit the cut off!!!!!
table(gi_sig, useNA="always")

starc_code_2$gi_sig = gi_sig # "Hot" or not
starc_code_2$Gi_value = gi_test_info$G.i # the actual Gi* value (not important)
starc_code_2$adj_zsc=gi_test_info$"Pr.z...E.G.i.." # the adjusted P-value

```


## Save STARC with Hot Spot Results Results
```{r}
st_write(starc_code_2, paste(folder, "/starc_code_gi.shp", sep=""), overwtite=T, append=F)

beep(10)

```



```{r}
# End banner when running App
if( exists("app_run")) {
removeModal()
}
```







