---
title: "Cluster Maps"
author: "Patrick Corbett"
date: "6/22/2023"
output: html_document
---


**Banner when running App**
```{r}
# Banner when running App
if( exists("app_run")) {
   showModal(modalDialog("Generating Clusters and Maps", footer=NULL))
}
```


**Set Parameters for running Outside the Rshiny App**
```{r, setup, include=FALSE}
time_start <- Sys.time()
if(! exists("app_run")) {
  #Establish Root Directory
  knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
  #Parameters
  countries="Burundi"
  }
```


# Additional Parameters
```{r}

buffer=500
```

# Setting Directory Names
```{r}
# countries="Malawi"
country_folder <- paste("Inputs/Country_Files/",countries,sep="")

output_folder <- paste0("Outputs/Country_Outputs/",countries )

folder <- paste0("Outputs/Country_Outputs/",countries, "/Clusters" )

dir.create(folder, recursive = TRUE)

```


##Load in the country boundary files

```{r}
ctry_border = st_read(paste0(country_folder,"/borders/admin_0_rgeo.shp" ))

# best CRS for this country
country_project = suggest_crs(ctry_border, gcs = 4326)[1,]
country_crs = country_project$crs_code

# transform the country border to best crs
ctry_border =st_transform(ctry_border, crs=as.numeric(country_crs))

# Read in the regions
regions = st_read(paste0(country_folder,"/borders/admin_1_rgeo.shp" ))
# Convert regions to best crs
regions =st_transform(regions, crs=as.numeric(country_crs))%>% dplyr::select(shapeName=shapeNm ,shapeGroup=shapGrp,shapeType=shapTyp ) #%>% dplyr::select(shapeName ,shapeGroup,shapeType ) #   


# Sub Regions
sub_region = st_read( paste0(country_folder,"/borders/admin_2_rgeo.shp"))
sub_region = st_transform(sub_region, crs(ctry_border))
# Recreate with both sub region and regional names
sub_regions2 = sub_region %>% st_join(regions %>% dplyr::select("Region"="shapeName"), largest=T) %>% dplyr::select(Region,"Sub_reg"="shapeNm")  #"Sub_reg"="shapeName") # 

```

## Load in STARC Files
```{r}
starc_code_2 = st_read(paste(folder, "/starc_code_gi.shp", sep=""))
hexes_cluster_final2 = st_read( paste(folder, "/hexes_in_cluster_final_", buffer,".shp", sep=""))
Cluster_outlines = st_read( paste(folder, "/cluster_outlines_", buffer,".shp", sep=""))

starc_unify = st_read(paste(output_folder, "/map_starc_unify.shp", sep=""))  

hot_spot_stats = st_read(paste(folder, "/hot_spot_outlines.shp", sep=""))

transmission_zones = st_read(paste(folder, "/zones_buffer_",buffer,".shp", sep="")) %>%
  mutate(level = factor(level, levels = c("Transmission","Susceptible"), labels = c("Transmission","Susceptible")))
transmission_clusters = st_read(paste(folder, "/zones_area_",buffer,".shp", sep="")) %>%
  mutate(level = factor(level, levels = c("Transmission","Susceptible"), labels = c("Transmission","Susceptible")))

```


## Load in Other Relevant Data
**These are used for mapping**
```{r}
# Unified STARC Map Created in create_map.RMD
#STARC_unify = st_read(paste0(output_folder, "/map_starc_unify.shp"))

# Open Street Map Layer (using library tmaptools)
osm_NLD <- tmaptools::read_osm(ctry_border, ext=1.1)

# Roads
roads <- st_read(paste0(country_folder,"/gis_osm_roads_free_1.shp" )) %>% filter(Road_class %in% c("Primary", "Secondary"))
roads <- st_transform(roads, crs(ctry_border))


```





# Map Clusters

## Create STARC_UNIFIED
**All other STARC Units NOT in the clusters...and not S6**
```{r}
All_Other_Hexes = starc_unify

```


## Assigning Color Variables
```{r}
# Assigning a ordered numeric ID to the STARC Codes. This is needed for tmap_polygons() to recognize the order of STARC codes when assigning pallette colors 
  # THere are many better ways to code this, but this works for now. 
strc_cd=sort(unique(starc_code_2$strc_cd))
colors=colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "Spectral"))(length(strc_cd))

starc_colors = data.frame(strc_cd,colors )

hexes_in_cluster <- hexes_cluster_final2 %>% merge(starc_colors, by="strc_cd", all.x=T)

All_Other_Hexes = starc_unify  %>% merge(starc_colors, by="strc_cd", all.x=T)




Hot_Spots = hot_spot_stats # to create borders of the hotspots outlined

prim_hexes = hexes_in_cluster %>% filter(Rep_pri=="primary")
sec_hexes = hexes_in_cluster %>% filter(Rep_pri=="secondary")
tert_hexes = hexes_in_cluster %>% filter(Rep_pri=="tertiary")
```


## Create PNG Map
```{r}
#tmap_options(check.and.fix = TRUE)
png_map <-tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
          tm_fill(col="white", alpha=0.6,popup.vars=FALSE)+
  tm_shape(regions) + 
          tm_borders(lwd =2) +
          #tm_fill( alpha=0,id="")+
  tm_shape(sub_regions2) + 
          tm_borders(lwd =0.5) +
          tm_fill( alpha=0,id="")+
  tm_shape(roads) +
          tm_lines(lwd=0.5, col="Road_class", palette=c("brown", "blue"),lty="dashed",legend.col.show=T, legend.lwd.show=F) + 
  tm_shape(Cluster_outlines) + 
          tm_polygons(a=0.6, lwd=0.1, id="cluster_id", col="Rep_pri", palette=c("red","blue", "pink")) + #,col="#000000" + 
  tm_shape(hexes_in_cluster) + 
          tm_polygons( border.alpha=0,
                      col="colors",title="STARC Codes", legend.hist=F, id="strc_cd") +
  tm_shape(Hot_Spots) + 
          tm_borders(alpha=0.9,lwd =0.4, col="black") +

  tm_layout(main.title=paste(countries,": Cluster MAP", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right")+ 
  tm_add_legend(type = "line", 
                col = c("black"),
                lwd = c(0.4),
                #lty = c(1, 4, 1),
                #labels = c("aaa", "bbb", "ccc"),
                title = "Hot Spots") +
  
  tm_add_legend(type = "fill", labels = strc_cd,col = colors,border.lwd = 0.5,title = "STARC Codes")+
  tm_compass(type = "8star", size = 4, position = c("right", "top"))+
  tm_scale_bar()

tmap_save(tm=png_map, filename= paste(folder, "/Clusters_",buffer,".png", sep=""), dpi=1300)

```

## PNG zone map
```{r}

reds = tmaptools::get_brewer_pal("YlOrRd", n = 5)

png_zone <-tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
  tm_shape(regions) + 
          tm_borders(lwd =2) +
  tm_shape(sub_regions2) + 
          tm_borders(lwd =0.5) +
  tm_shape(roads) +
          tm_lines(lwd=0.5, col="Road_class", palette=c("dodgerblue4", "dodgerblue"),lty="dashed",legend.col.show=T, legend.lwd.show=F) + 
  tm_shape(transmission_zones) + tm_fill(col = reds[1], alpha = 1) + tm_borders(lwd = .3, col = "black") + 
  tm_shape(transmission_clusters) + tm_fill(col = reds[3], alpha = 1) +
  tm_shape(Hot_Spots) + tm_fill(col = reds[5], alpha = 1) +
  tm_add_legend(type = "fill", col = c(reds[c(1,3,5)]), labels = c("Transmission Zones","Clusters","Hot Spots"), border.alpha = 0, title = "Risk Area") +
  tm_layout(main.title=paste(countries,": Risk Map", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right")+ 
  # tm_add_legend(type = "fill", labels = strc_cd,col = colors,border.lwd = 0.5,title = "STARC Codes")+
  tm_compass(type = "8star", size = 4, position = c("right", "top"))+
  tm_scale_bar()


tmap_save(tm=png_zone, filename= paste(folder, "/Zone_Map_",buffer,".png", sep=""), dpi=1300)

```

## PNG risk map
```{r}

png_risk <- tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
  tm_shape(regions) + 
          tm_borders(lwd =2) +
  tm_shape(sub_regions2) +
          tm_borders(lwd =0.5) +
  tm_shape(transmission_zones) + tm_fill(col = "level", alpha = .5, palette = c("#BE1826","#3A55A4")) + 
  # tm_borders(lwd = .3, col = "black") + 
  tm_shape(transmission_clusters) + tm_fill(col = "level", alpha = .5, palette = c("#BE1826","#3A55A4"),
                                            legend.show = FALSE) +
  # tm_shape(Hot_Spots) + tm_fill(col = reds[5], alpha = .7) +
  tm_layout(main.title=paste(countries,": Risk Map", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right")+ 
  tm_compass(type = "8star", size = 2, position = c("right", "top"))+
  tm_scale_bar()


tmap_save(tm=png_risk, filename= paste(folder, "/Risk_Map_",buffer,".png", sep=""), dpi=1300)



```


## Create HTML Map
```{r}

html_map <-tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
          tm_fill(col="white", alpha=0.6,popup.vars=FALSE)+
  tm_shape(regions) + 
          tm_borders(lwd =2) +
  tm_shape(sub_regions2) + 
          tm_borders(lwd =0.5) +
          tm_fill( alpha=0,id="")+
 
  tm_shape(All_Other_Hexes) + 
          tm_polygons(col="colors", border.alpha=0,title="ALL Other STARC Units", legend.hist=F, id="strc_cd") +
  tm_shape(Cluster_outlines) + 
          tm_polygons(a=0.6, lwd=0.1, id="cluster_id", title="Cluster Outlines",col="Rep_pri", palette=c("red","blue", "pink"),popup.vars=colnames(Cluster_outlines)[colnames(Cluster_outlines)!="geometry"]) + 
  
  tm_shape(prim_hexes) + 
          tm_polygons( border.alpha=0,
                      col="colors",title="Primary Cluster", legend.hist=F, id="strc_cd") 

if(nrow(sec_hexes)>0){
  html_map=html_map+tm_shape(sec_hexes) + 
          tm_polygons( border.alpha=0,
                      col="colors",title="Secondary Cluster", legend.hist=F, id="strc_cd")
}
if(nrow(tert_hexes)>0){
  html_map=html_map+tm_shape(tert_hexes) + 
          tm_polygons( border.alpha=0,
                      col="colors",title="Tertiary Cluster", legend.hist=F, id="strc_cd") 
}

html_map = html_map +
  tm_shape(Hot_Spots) + 
          tm_borders(alpha=0.9,lwd =0.4, col="black") +
   tm_shape(roads) +
          tm_lines(lwd=0.5, col="Road_class", palette=c("brown", "blue"),lty="dashed",legend.col.show=T, legend.lwd.show=F) +
  

  tm_layout(main.title=paste(countries,": Cluster MAP", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right")+ 
  tm_add_legend(type = "line", 
              col = "black",
              lwd = 0.4,
              #lty = c(1, 4, 1),
              #labels = c("aaa", "bbb", "ccc"),
              title = "Hot Spots") +
  tm_add_legend(type = "fill", labels = strc_cd,col = colors,border.lwd = 0.5,title = "STARC Codes")+
  tm_compass(type = "8star", size = 4, position = c("right", "top"))+
  tm_scale_bar()

```


```{r}
tmap_options(check.and.fix = TRUE)
html_map2=  addDrawToolbar(map= tmap_leaflet(html_map))%>%
  hideGroup("All_Other_Hexes") #%>%
  #addMeasure()# %>% # https://rstudio.github.io/leaflet/morefeatures.html
  #addResetMapButton() %>% #https://www.r-bloggers.com/2019/12/quick-tips-for-customizing-your-r-leaflet-map/
  #addSearchFeatures(targetGroups=Cluster_outlines) #https://cran.r-project.org/web/packages/leaflet.extras/leaflet.extras.pdf

saveWidget(html_map2, file= paste(folder, "/Cluster2_",buffer,".html", sep=""), selfcontained=F)


```

## HTML zone map
```{r}

html_zone <-tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
  tm_shape(regions) + 
          tm_borders(lwd =2) +
  tm_shape(sub_regions2) + 
          tm_borders(lwd =0.5) +
  tm_shape(All_Other_Hexes) + 
          tm_polygons(col="colors", border.alpha=0,title="ALL Other STARC Units", legend.hist=F, id="strc_cd", alpha = .6) +
  tm_shape(hexes_in_cluster) + 
          tm_polygons( border.alpha=0, alpha = .6,
                      col="colors",title="Primary Cluster", legend.hist=F, id="strc_cd") +
  tm_shape(transmission_zones) + tm_fill(col = reds[1], alpha = .8) + tm_borders(lwd = .3, col = "black") + 
  tm_shape(transmission_clusters) + tm_fill(col = reds[3], alpha = .8) +
  tm_shape(Hot_Spots) + tm_fill(col = reds[5], alpha = .8) +
  tm_add_legend(type = "fill", col = c(reds[c(1,3,5)]), labels = c("Transmission Zones","Clusters","Hot Spots"),
                border.alpha = 0, title = "Risk Area") +
   tm_shape(roads) +
          tm_lines(lwd=0.5, col="Road_class", palette=c("dodgerblue4", "dodgerblue"),lty="dashed",legend.col.show=T, legend.lwd.show=F) +
  tm_layout(main.title=paste(countries,": Risk Map", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right") + 
  tm_add_legend(type = "fill", labels = strc_cd,col = colors,border.lwd = 0.5,title = "STARC Codes")+
  tm_compass(type = "8star", size = 4, position = c("right", "top"))+
  tm_scale_bar()

# tmap_mode("plot")
# html_map

```
```{r}
html_zone2=  addDrawToolbar(map= tmap_leaflet(html_zone)) %>%
  hideGroup(c("All_Other_Hexes","hexes_in_cluster")) #%>%
  #addMeasure()# %>% # https://rstudio.github.io/leaflet/morefeatures.html
  #addResetMapButton() %>% #https://www.r-bloggers.com/2019/12/quick-tips-for-customizing-your-r-leaflet-map/
  #addSearchFeatures(targetGroups=Cluster_outlines) #https://cran.r-project.org/web/packages/leaflet.extras/leaflet.extras.pdf

saveWidget(html_zone2, file= paste(folder, "/Zone_Map_",buffer,".html", sep=""), selfcontained=F)


```

## HTML risk map
```{r}

html_risk <-  tm_shape(osm_NLD) + tm_rgb()+
  tm_shape(ctry_border) + 
          tm_borders(lwd =1) + 
  tm_shape(regions) + 
          tm_borders(lwd =2) +
  tm_shape(sub_regions2) +
          tm_borders(lwd =0.5) +
  tm_shape(All_Other_Hexes) +
          tm_polygons(col="colors", border.alpha=0,title="ALL Other STARC Units", legend.hist=F, id="strc_cd", alpha = .6) +
  tm_shape(hexes_in_cluster) +
          tm_polygons( border.alpha=0, alpha = .6,
                      col="colors",title="Primary Cluster", legend.hist=F, id="strc_cd") +
  tm_shape(transmission_zones) + tm_fill(col = "level", alpha = .5, palette = c("#BE1826","#3A55A4")) + 
  # tm_borders(lwd = .3, col = "black") + 
  tm_shape(transmission_clusters) + tm_fill(col = "level", alpha = .5, palette = c("#BE1826","#3A55A4"),
                                            legend.show = FALSE) +
  # tm_shape(Hot_Spots) + tm_fill(col = reds[5], alpha = .7) +
  # tm_add_legend(type = "fill", col = c(reds[c(1,3,5)]), labels = c("Transmission Zones","Clusters","Hot Spots"),
                # border.alpha = 0, title = "Risk Area") +
   tm_shape(roads) +
          tm_lines(lwd=0.5, col="Road_class", palette=c("dodgerblue4", "dodgerblue"),lty="dashed",legend.col.show=T, legend.lwd.show=F) +
  tm_layout(main.title=paste(countries,": Risk Map", sep=""), main.title.position = c('center', 'top'), legend.outside = TRUE, legend.outside.position = "right") + 
  tm_add_legend(type = "fill", labels = strc_cd,col = colors,border.lwd = 0.5,title = "STARC Codes")+
  tm_compass(type = "8star", size = 4, position = c("right", "top"))+
  tm_scale_bar()

```

```{r}
html_risk2=  addDrawToolbar(map= tmap_leaflet(html_risk)) %>%
  hideGroup(c("All_Other_Hexes","hexes_in_cluster")) #%>%
  #addMeasure()# %>% # https://rstudio.github.io/leaflet/morefeatures.html
  #addResetMapButton() %>% #https://www.r-bloggers.com/2019/12/quick-tips-for-customizing-your-r-leaflet-map/
  #addSearchFeatures(targetGroups=Cluster_outlines) #https://cran.r-project.org/web/packages/leaflet.extras/leaflet.extras.pdf

saveWidget(html_risk2, file= paste(folder, "/Risk_Map_",buffer,".html", sep=""), selfcontained=F)


```


## SAVE Maps
```{r}
#tmap_save(tm=png_map, filename= paste(folder, "/Clusters_",buffer,".png", sep=""), dpi=1300)

#tmap_save(tm=html_map, filename= paste(folder, "/Clusters2_",buffer,".html", sep=""))

# library(mapview)
# mapview::mapshot(html_map2, file= paste(folder, "/Cluster_",buffer,".html", sep=""))

beep(2)

```

```{r}
# End banner when running App
if( exists("app_run")) {
removeModal()
}
```
